---
title: "Country differences in ERC success rates"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message = FALSE,warning = FALSE)

library(tidyverse)
library(readxl)
library(eurostat)
library(countrycode)
library(ggplot2)
library(ggrepel)
library(patchwork)

```
 
### Rui Mata, [University of Basel](https://www.cds.unibas.ch)

## Goal 

I was interested in assessing country differences in ERC success rates and potential relation to investment in research and development, with a particular focus on Portugal, Germany, and Switzerland. 

## Data

First, I manually downloaded publicly available statiscts from [erc.europa.eu](https://erc.europa.eu/projects-figures/statistics) that included the number of evaluated and granted proposals for ERC Starting, Consolidator, and Advanced Grants. 

```{r}

starting_eval <- read_excel("starting_eval.xlsx")
consolidator_eval <- read_excel("consolidator_eval.xlsx")
advanced_eval <- read_excel("advanced_eval.xlsx")
starting_granted <- read_excel("starting_granted.xlsx")
consolidator_granted <- read_excel("consolidator_granted.xlsx")
advanced_granted <- read_excel("advanced_granted.xlsx")

starting_eval <- starting_eval %>% 
  mutate(cat="starting") %>% 
  pivot_longer("2007":"2021",names_to = "year",values_to = "number")

consolidator_eval <- consolidator_eval %>% 
  mutate(cat="consolidator") %>% 
  pivot_longer("2013":"2021",names_to = "year",values_to = "number")

advanced_eval <- advanced_eval %>% 
  mutate(cat="advanced") %>% 
  pivot_longer("2008":"2020",names_to = "year",values_to = "number")

starting_granted <- starting_granted %>% 
  mutate(cat="starting") %>% 
  pivot_longer("2007":"2021",names_to = "year",values_to = "number")

consolidator_granted <- consolidator_granted %>% 
  mutate(cat="consolidator") %>% 
  pivot_longer("2013":"2021",names_to = "year",values_to = "number")

advanced_granted <- advanced_granted %>% 
  mutate(cat="advanced") %>% 
  pivot_longer("2008":"2020",names_to = "year",values_to = "number")

starting<-left_join(starting_eval,starting_granted,by=c("Country","year","cat")) %>% 
  rename(applied=number.x, 
         granted=number.y) %>% 
  mutate(granted=if_else(!is.na(applied) & is.na(granted),0,granted)) %>% 
  mutate(rate=granted/applied*100) %>% 
  mutate(year=as.Date(year,format='%Y')) %>% 
  filter(!is.na(applied)) 

consolidator<-left_join(consolidator_eval,consolidator_granted,by=c("Country","year","cat")) %>% 
  rename(applied=number.x, 
         granted=number.y) %>% 
  mutate(granted=if_else(!is.na(applied) & is.na(granted),0,granted)) %>% 
  mutate(rate=granted/applied*100) %>% 
  mutate(year=as.Date(year,format='%Y')) %>% 
  filter(!is.na(applied)) 

advanced<-left_join(advanced_eval,advanced_granted,by=c("Country","year","cat")) %>% 
  rename(applied=number.x, 
         granted=number.y) %>% 
  mutate(granted=if_else(!is.na(applied) & is.na(granted),0,granted)) %>% 
  mutate(rate=granted/applied*100) %>% 
  mutate(year=as.Date(year,format='%Y')) %>% 
  filter(!is.na(applied)) 

erc<-rbind(starting,consolidator,advanced) 

erc<-erc %>% 
  mutate(iso=countrycode(Country,origin="country.name",destination = 'iso2c'))
```

Second, I used the eurostat R package to access [eurostat](https://ec.europa.eu/eurostat/statistics-explained/index.php?title=R%26D_expenditure&oldid=551418#Gross_domestic_expenditure_on_R.26D) statistics concerning the Gross domestic Expenditure on Research and Development (GERD) of European countries. 

```{r}
# get GDP data

# check out: https://ec.europa.eu/eurostat/statistics-explained/index.php?title=R%26D_expenditure&oldid=551418#Gross_domestic_expenditure_on_R.26D
#data=get_eurostat("rd_e_gerdtot",time_format = "num")
#write_csv(data,"gdp.csv")

#data=get_eurostat("tsc00007",time_format = "num")
#write_csv(data,"gdp.csv")

gdp=read_csv("gdp.csv")

gdp <- gdp %>% 
  mutate(iso=countrycode(geo,origin="eurostat",destination = 'iso2c')) %>% 
    filter(time>2000,
         unit=="PC_GEXP",
         !is.na(iso)) %>% 
  group_by(iso) %>% 
  summarize(gdp=mean(values)) %>% 
  ungroup()

```

## Results


```{r}

success= sum(erc$granted)/sum(erc$applied)*100

summary_table <- erc %>% 
  group_by(iso) %>% 
  summarize(rate=median(rate)) %>% 
  mutate(Success=round(rate)) %>% 
  mutate(Country=countrycode(iso,origin='iso2c',destination="country.name")) %>% 
  #filter(Country=="Portugal" | Country=="Germany" | Country=="Switzerland") %>% 
  select(Country,Success)

```
#### Overall Success Rate

I calculated the overall success rate across all available years for Starting, Consolidator, and Advanced Grants as the total number of funded grants divided by total submissions. 

The overall rate was `r round(success,1)`% but this global rate hides considerable country differences. The table below shows success rates, including those for PT (`r summary_table[summary_table$Country=="Portugal" & !is.na(summary_table$Country=="Portugal"),"Success"]`%), DE (`r summary_table[summary_table$Country=="Germany" & !is.na(summary_table$Country=="Germany"),"Success"]`%), and CH (`r summary_table[summary_table$Country=="Switzerland" & !is.na(summary_table$Country=="Switzerland"),"Success"]`%). 

```{r}



DT::datatable(summary_table)

```

#### Success Rate by GERD

To explore the link between country differences in ERC success rates and each country's investment in research and development, I used a scatterplot linking success rates as calculated above and the GERD (Gross domestic Expenditure on Research and Development) data available from Eurostat. 

```{r}

summary_erc <- erc %>% 
  group_by(iso) %>% 
  summarize(rate=median(rate)) 

summary_gdp <- gdp %>% 
  group_by(iso) %>% 
  summarize(gdp=median(gdp)) %>% 
  filter(!is.na(gdp))

data=left_join(summary_erc,summary_gdp,by="iso")

selection_data <- data %>% 
  filter(iso=="PT" | iso=="DE" | iso=="CH")

ggplot(selection_data,aes(x=gdp,y=rate,label=iso)) +
  geom_point(alpha=.4) +
  geom_text_repel(size=5,min.segment.length = .25) +
  theme_minimal() +
  xlab("R&D (% GDP)") +
  ylab("Success Rate (%)") +
  theme(title=element_text(size=25),axis.title = element_text(size=20),axis.text = element_text(size=15)) +
  xlim(0,3) +
  ylim(0,25)

ggsave("success_gdp_selection.jpeg", width = 6, height = 4)
```

```{r}

ggplot(data,aes(x=gdp,y=rate,label=iso)) +
  geom_point(alpha=.4) +
  geom_text_repel(size=5,min.segment.length = .25) +
  theme_minimal() +
  xlab("R&D (% GDP)") +
  ylab("Success Rate (%)") +
  theme(title=element_text(size=25),axis.title = element_text(size=20),axis.text = element_text(size=15)) +
  xlim(0,3) +
  ylim(0,25)

ggsave("success_gdp.jpeg", width = 6, height = 4)
```
